Whose voices are heard?

### 2024-02-29 paper/app-eventdata.qmd
re-submitted to IJPP

The first event-level variable, *Protest Goal*, originally contained eight different demands from protesters.
If a protest had more than one demand, @ClarkMMPdata2019 coded multiple goals for the same protest accordingly.
The specific goals coded in the MMP data were recoded to achieve a better fit with the literature reviewed in Section \@ref(sec:factors).
Table \@ref(tab:goalsrecode) shows the codes used by @ClarkMMPdata2019 on the left and the respective codes used here on the right.^[
Specifically, the categories used here are *anti-war*, *labour*, *police brutality*, *social-issue* and *political* protests.
]
To obtain the *anti-war* coding, protester identity, which is another code used in the MMP data, was used in addition to the protest demand.
Table \@ref(tab:goalsdist) shows how often each demand was coded.

```{r goalsrecode}
tibble::tribble(
  ~Code, ~"Recoded to",
  "political behavior, process", "anti-war (also based on protester identity)",
  "labor wage dispute", "labour",
  "police brutality", "police brutality",
  "land farm issue", "social-issue",
  "price increases, tax policy", "social-issue",
  "social restriction", "social-issue",
  "political behavior, process", "political",
  "removal of politician", "political"
) %>%
  custom_df_print(caption = "Recoded Protester Demands")
```

```{r goalsdist}
rio::import("../misc/protesterdemand_table.csv") |>
  custom_df_print(caption = "Distribution of Protester Demands")
```

@@

As mentioned in Section \@ref(sec:eventdata), a protest's goal and violence are prone to be affected by endogeneity, as @ClarkMMPdata2019 coded their data based on news reports, which is also the basis for the dependent variable here.
It needs to be noted that the reports used for coding were removed before calculating the models.^[
Specifically, the MMP data contains a text snippet from the article(s) on which the coding was based. These articles were removed from the database. Additional robustness checks were also performed and are described later on.
]
Nevertheless, the results regarding the variables *protest goal* and *violence* should be seen with some caution.

The variable *state repression of peaceful protest* was constructed to reflect the occurrence of a specific case outlined by @WasowSeeding2020:
when the state uses repression against peaceful protesters.
In this case, the expectation is that there will be more legitimising framing in the coverage of a protest event.
The variable *protester violence* was used again to assess the second condition for this case.
For the first condition, the variable *stateresponse* in the MMP dataset was employed.
Of the five coded responses, "arrests" or "beatings" were interpreted as being repressive.
When one of these two was employed by the state during peaceful protests, *state repression of peaceful protest* was coded as 1, otherwise, it was 0 ($\bar{x} = 0.46$).
Figure \@ref(fig:stateresponse) shows the percentage of violent and peaceful protests facing different kinds of state responses in the subset of the MMP data.
One problem becomes apparent in this:
the data for the UK did not contain a single case in which beatings occurred during peaceful protests.
Therefore, in this case, *state repression of peaceful protest* reflects arrests of peaceful protesters.
This is specific to the UK, data whereas beatings of peaceful protesters were recorded in other countries, including other countries in Western Europe.

```{r stateresponse, fig.height=3, fig.width=6, fig.cap="Percent of Violent and Peaceful Protests Facing Different Kinds of State Responses"}
rio::import("../misc/protesterviolence_table.csv") %>%
  ggplot(aes(x = pct, y = as.factor(protesterviolence), fill = stateresponse2)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  scale_fill_jbg() +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none") +
  annotate(geom = "text", x = 0.10, y = 2.9, label = "ignore",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.105, y = 2.9, xend = 0.2, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.4, y = 2.9, label = "accomodation",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.405, y = 2.9, xend = 0.5, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.68, y = 2.9, label = "crowd dispersal",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.685, y = 2.9, xend = 0.71, yend = 2.5,
    curvature = -.2, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.87, y = 2.9, label = "arrests",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.875, y = 2.9, xend = 0.93, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.9, y = 0.2, label = "beatings",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.905, y = 0.2, xend = 0.95, yend = 0.5,
    curvature = .3, colour = "#606367"
  ) +
  coord_cartesian(clip = 'off')
```

The three outlet-level variables were coded based on the literature on the characteristics of the UK newspapers.
Specifically, newspaper outlets in the UK are divided into right-leaning (i.e., closer to the Conservative party) and left-leaning (i.e., closer to the Labour party), as well as between tabloids and broadsheets.
The newspaper outlets included in the sample were coded as noted in Table \@ref(tab:outlets).

```{r}
ideo_count_table <- rio::import("../misc/ideo_count_table.csv")
ideo <- function(l, type = "pct") {
  ideo_count_table %>%
    filter(ideology == l) %>%
    select(matches(type)) %>%
    pull()
}
```

@@

The table below contains the ideology variable that was added through annotation of events.

@@

```{r}
rio::import("../misc/mmp_events_coded_table.csv") %>%
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "",
               caption = "Employed Subset of the MMP Data by Clark and Regan (2019) with Additional and Recoded Variables",
               label = NA) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    latex_options = "repeat_header",
    font_size = 10
  )%>% 
  kableExtra::column_spec(c(1:3, 5), width = "1.8cm") %>%
  kableExtra::column_spec(4, width = "6cm")
```

>>>>

The first event-level variable, *Protest Goal*, originally contained eight different demands from protesters.
If a protest had more than one demand, @ClarkMMPdata2019 coded multiple goals for the same protest accordingly.
The specific goals coded in the MMP data were recoded to achieve a better fit with the literature reviewed in Section *Event-Level Factors*.
@tbl-goalsrecode shows the codes used by @ClarkMMPdata2019 on the left and the respective codes used here on the right.^[
Specifically, the categories used here are *anti-war*, *labour*, *police brutality*, *social-issue* and *political* protests.
]
To obtain the *anti-war* coding, protester identity, which is another code used in the MMP data, was used in addition to the protest demand.
@tbl-goalsrecode shows how often each demand was coded.

```{r tbl-goalsrecode}
#| tbl-cap: Recoded Protester Demands
tibble::tribble(
  ~Code, ~"Recoded to",
  "political behavior, process", "anti-war (also based on protester identity)",
  "labor wage dispute", "labour",
  "police brutality", "police brutality",
  "land farm issue", "social-issue",
  "price increases, tax policy", "social-issue",
  "social restriction", "social-issue",
  "political behavior, process", "political",
  "removal of politician", "political"
) |> 
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "")
```

```{r tbl-goalsdist}
#| tbl-cap: Distribution of Protester Demands
# rio::import("../misc/protesterdemand_table.csv") |> 
tibble::tribble(
  ~Demand, ~n, ~"of total",
  "political", 3656, "64.8%",
  "social-issue", 1102, "19.5%",
  "labour", 566, "10.0%",
  "anti-war", 225, "4.0%",
  "police", 90, "1.6%"
) |> 
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "")
  # custom_df_print(caption = "Distribution of Protester Demands")
```

@@

The variable *state repression of peaceful protest* was constructed to reflect the occurrence of a specific case outlined by @WasowSeeding2020:
when the state uses repression against peaceful protesters.
In this case, the expectation is that there will be more legitimising framing in the coverage of a protest event.
The variable *protester violence* was used again to assess the second condition for this case.
For the first condition, the variable *stateresponse* in the MMP dataset was employed.
Of the five coded responses, "arrests" or "beatings" were interpreted as being repressive.
When one of these two was employed by the state during peaceful protests, *state repression of peaceful protest* was coded as 1, otherwise, it was 0 ($\bar{x} = 0.46$).
@fig-stateresponse shows the percentage of violent and peaceful protests facing different kinds of state responses in the subset of the MMP data.
One problem becomes apparent in this:
the data for the UK did not contain a single case in which beatings occurred during peaceful protests.
Therefore, in this case, *state repression of peaceful protest* reflects arrests of peaceful protesters.
This is specific to the UK, data whereas beatings of peaceful protesters were recorded in other countries, including other countries in Western Europe.

```{r fig-stateresponse, fig.height=3, fig.width=6, fig.cap="Percent of Violent and Peaceful Protests Facing Different Kinds of State Responses"}
rio::import("../misc/protesterviolence_table.csv") |>
  ggplot(aes(x = pct, y = as.factor(protesterviolence), fill = stateresponse2)) +
  geom_col() +
  scale_x_continuous(labels = scales::percent) +
  scale_fill_jbg() +
  labs(x = NULL, y = NULL, fill = NULL) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position = "none") +
  annotate(geom = "text", x = 0.10, y = 2.9, label = "ignore",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.105, y = 2.9, xend = 0.2, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.4, y = 2.9, label = "accomodation",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.405, y = 2.9, xend = 0.5, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.68, y = 2.9, label = "crowd dispersal",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.685, y = 2.9, xend = 0.71, yend = 2.5,
    curvature = -.2, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.87, y = 2.9, label = "arrests",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.875, y = 2.9, xend = 0.93, yend = 2.5,
    curvature = -.3, colour = "#606367"
  ) +
  annotate(geom = "text", x = 0.9, y = 0.2, label = "beatings",
           hjust = "right", colour = "#606367", family = "Times") +
  annotate(
    geom = "curve", x = 0.905, y = 0.2, xend = 0.95, yend = 0.5,
    curvature = .3, colour = "#606367"
  ) +
  coord_cartesian(clip = 'off')
```

The three outlet-level variables were coded based on the literature on the characteristics of the UK newspapers.
Specifically, newspaper outlets in the UK are divided into right-leaning (i.e., closer to the Conservative party) and left-leaning (i.e., closer to the Labour party), as well as between tabloids and broadsheets.
The newspaper outlets included in the sample were coded as noted in @tbl-outlets.

```{r}
ideo_count_table <- rio::import("../misc/ideo_count_table.csv")
ideo <- function(l, type = "pct") {
  ideo_count_table |>
    filter(ideology == l) |>
    select(matches(type)) |>
    pull()
}
```

@@

@tbl-eventdata contains the ideology variable that was added through annotation of events.

@@

```{r tbl-eventdata}
#| eval: true
#| tbl-cap: Employed Subset of the MMP Data by Clark and Regan (2019) with Additional and Recoded Variables
rio::import("../misc/mmp_events_coded_table.csv") |> 
  knitr::kable(format = "latex",
               booktabs = TRUE,
               longtable = TRUE,
               linesep = "") |>
  kableExtra::kable_styling(
    full_width = FALSE,
    latex_options = "repeat_header",
    font_size = 10
  )|> 
  kableExtra::column_spec(c(1:3, 5), width = "1.8cm") |>
  kableExtra::column_spec(4, width = "6cm")
```

### 2024-02-29 paper/article.qmd
re-submitted to IJPP

format:
  aft-pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
date: today
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
bibliography: references.bib
execute:
  echo: false
  eval: true
---

@@

# What Drives Protest Coverage in News Media? {#sec:factors}

@@

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from @gruberTrouble2023.
In short, frames were identified by coding the frame elements found in Entman's [-@entman1993] definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles.
These elements were then combined to frames using factor analysis.
The result is one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models on the manually coded random sample, all remaining articles were coded using these models.

@@

I merged the two datasets by using the time frame of each event and some keywords (see Appendix).
Overall, 5,639 articles could be matched with the 471 events from the database.
Using the variables listed in @tbl-variables, I then fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectively constitute a panel dataset, since observations must be seen as nested and not independent.
<!-- I used newspaper-year as the Level 2 variable ($n = 180$), since the number of newspapers alone was considered too low for this design. -->
<!-- It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets. -->
<!-- However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016]. -->
<!-- Therefore, I used newspaper-year as the Level 2 variable ($n = 180$). -->
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->
<!-- I assessed whether adding the newspaper-year level increases or decreases model fit. -->
<!-- To compare models, the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) are calculated and compared using Analysis Of Variance (ANOVA) tests. -->
<!-- The lower the AIC and BIC, the better the model fits the data. -->
<!-- Both AIC and BIC are always considerably lower when including the second level into the modelling. -->
<!-- The p-values suggest that including the random effect improves all models significantly. -->
<!-- I examined if adding a newspaper-year level improves model fit. -->
<!-- The model comparisons were done using AIC, BIC, and ANOVA tests. -->
The inclusion of the second level significantly lowers both AIC and BIC for all individual and grouped frame models.
<!-- , and the p-values indicate a significant improvement in all models. -->

@@

As was already shown in @gruberTrouble2023, the year does not affect the probability of delegitimising frames but does increase the probability that legitimising frames being used.
<!-- ^[ -->
<!-- Although @gruberTrouble2023 uses the year in which an article was published while here it is the year of the protest event itself --- which is identical for most articles though. Note also that articles were excluded if they could not be matched with any event in the MMP data. -->
<!-- ] -->
The model confirms this previous finding as the year has a significant positive effect on the probability that legitimising framing is used, suggesting legitimising frames become more salient over time.
Specifically, comparing the last year, 2017, in the dataset to the first year, 1992, legitimising framing has become `r marg("Year of protest (since 1992)", model = "legitimising", unit = "percent")` more likely on average.
H8b is this confirmed.

@@

# Appendix {#sec-eventdata .unnumbered}

>>>>

format:
  aft-pdf:
    keep-tex: true
    include-in-header:
      - text: |
          \usepackage{siunitx}
          \usepackage{graphicx}
          \usepackage{rotating}
          \usepackage{setspace}
          \linespread{2}
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
  aft-html: 
    citeproc: false
    filters: 
      - "_extensions/andrewheiss/wordcount/citeproc.lua"
      - "_extensions/andrewheiss/wordcount/wordcount.lua"
filters:
  - abstract-section
keywords:
    - "protest paradigm"
    - "media framing"
    - "social movements"
    - "news reporting"
    - "content analysis"
bibliography: references.bib
execute:
  echo: false
  eval: true
---

@@

# What Drives Protest Coverage in News Media? {#sec-factors}

@@

This dataset was coded for framing of protest reporting.
Specifically, I re-use the measurement of framing in newspaper reports on protests and demonstrations from Author (Year). <!-- @gruberTrouble2023. -->
In short, frames were identified by coding the frame elements found in Entman's [-@entman1993] definition of framing (problem definition, causal interpretation, moral evaluation, and treatment recommendation) in 500 randomly selected articles.
These elements were then combined to frames using factor analysis.
The result is one binary variable per frame, which takes the value 1 if a frame is present and 0 otherwise.
Each article can contain several frames, or none.
After training supervised machine learning models on the manually coded random sample, all remaining articles were coded using these models.

@@

I merged the two datasets by using the time frame of each event and some keywords (see Supplementary Information file).
Overall, 5,639 articles could be matched with the 471 events from the database.
Using the variables listed in @tbl-variables, I then fitted multilevel logistic regression models (one for each frame) to establish the influence of the different factors on reporting.
Specifically, I used the *lme4* package in R [@lme4].
This is appropriate as the dependent variables for all models are binary and the articles published in the same newspaper and in the same year effectively constitute a panel dataset, since observations must be seen as nested and not independent.
<!-- I used newspaper-year as the Level 2 variable ($n = 180$), since the number of newspapers alone was considered too low for this design. -->
<!-- It was considered to use newspapers as the second level variable as unobserved and unobservable heterogeneity is most likely present between outlets. -->
<!-- However, the number of newspapers in the dataset (eight) is considered problematically low for fitting a valid multilevel model [e.g., @bryanmultilevel2016]. -->
<!-- Therefore, I used newspaper-year as the Level 2 variable ($n = 180$). -->
<!-- That means that *The Guardian*, for example, is treated as a different second-level case in 1992, 1993, 1994 and so on.
This makes sense as the last chapter has shown that there is a difference in reporting between all newspaper over time. -->
<!-- I assessed whether adding the newspaper-year level increases or decreases model fit. -->
<!-- To compare models, the Akaike information criterion (AIC) and the Bayesian information criterion (BIC) are calculated and compared using Analysis Of Variance (ANOVA) tests. -->
<!-- The lower the AIC and BIC, the better the model fits the data. -->
<!-- Both AIC and BIC are always considerably lower when including the second level into the modelling. -->
<!-- The p-values suggest that including the random effect improves all models significantly. -->
<!-- I examined if adding a newspaper-year level improves model fit. -->
<!-- The model comparisons were done using AIC, BIC, and ANOVA tests. -->
The inclusion of the second level significantly lowers both AIC and BIC for all individual and grouped frame models.
<!-- , and the p-values indicate a significant improvement in all models. -->

@@

<!-- As was already shown in  -->
@gruberTrouble2023 showed that the year does not affect the probability of delegitimising frames but does increase the probability that legitimising frames being used.
<!-- ^[ -->
<!-- Although @gruberTrouble2023 uses the year in which an article was published while here it is the year of the protest event itself --- which is identical for most articles though. Note also that articles were excluded if they could not be matched with any event in the MMP data. -->
<!-- ] -->
The model confirms this previous finding as the year has a significant positive effect on the probability that legitimising framing is used, suggesting legitimising frames become more salient over time.
Specifically, comparing the last year, 2017, in the dataset to the first year, 1992, legitimising framing has become `r marg("Year of protest (since 1992)", model = "legitimising", unit = "percent")` more likely on average.
H8b is this confirmed.

@@

# Supplementary Information file {#sec-eventdata .unnumbered}

\newpage
\appendix
\renewcommand{\thefigure}{A\arabic{figure}}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
